---
title: "notebook2"
output: html_document
date: "2024-06-09"
---

First, let's download the required packages.
```{r}
if (!requireNamespace("randomForest", quietly = TRUE)) {
  install.packages("randomForest")
}
library(randomForest)
```

Let's set the working directory and load the training and test data. 
```{r}
setwd('/Users/fredrikwhaug/code/bst-209-notebooks')
config <- config::get()
data_train<- read.csv(file.path(config$data_split, 'wids_train.csv'))
data_test <- read.csv(file.path(config$data_split, 'wids_test.csv'))
```

Let's look at the dimensions of our traning and test data.
```{r}
dim(data_train)
dim(data_test)
```
Define the outcome column and extract dependend variables. 
```{r}
outcome <- data$hospital_death
y_train <- data_train$outcome
y_test <- data_test$outcome
```

Select the features to use. For now, we will use the original SpO2 and Lactate 
features as defined in the dataset. Please define which other features you, as a
team, want to use!
```{r}
features <- data[, c(
    'd1_spo2_min',
    'd1_lactate_max',
    'bmi',
    'age',
    'ethnicity',
    'gender'
    # ...
)]
```

Define the independent variables.
```{r}
X_train <- data_train$features
X_test <- data_test$features
```

```{r}
# Fit the model
#model <- randomForest(X_train, y_train)

# Make predictions
#predictions <- predict(model, X_test)

# Predict probabilities
#prob_predictions <- predict(model, X_test, type = "prob")

# Display predictions and probabilities
#print(predictions)
#print(prob_predictions)
```

Evaluate your model across race and ethnicity.
```{r}
# first split the y_test per racial and ethnic group
# then for each group compute the evaluation metrics
# and you can compute overall metrics as well!
```