---
title: "notebook3"
output: html_document
date: "2024-06-09"
---

```{r}
# Some possibilities for modeling and evaluation, uncomment if necessary!

# Load necessary libraries
# install.packages("glmnet") # for logistic regression
# install.packages("randomForest") # for random forest
# install.packages("Metrics") # for evaluation metrics

# Load libraries
# library(glmnet)
# library(randomForest)
# library(Metrics)

# Logistic Regression
# model <- glmnet(as.matrix(X_train), y_train, family = "binomial")

# Random Forest Classifier
# model <- randomForest(X_train, as.factor(y_train))

# Predictions
# predictions <- predict(model, as.matrix(X_test), type = "response")
# if using random forest, use type = "class" for class predictions
# predictions <- predict(model, X_test, type = "class")

# Predict probabilities
# prob_predictions <- predict(model, as.matrix(X_test), type = "response")
# if using random forest, use type = "prob" for probability predictions
# prob_predictions <- predict(model, X_test, type = "prob")

# Evaluation Metrics
# roc_auc <- auc(y_test, prob_predictions)
# avg_precision <- mean(average_precision_score(y_test, prob_predictions))
# accuracy <- accuracy_score(y_test, predictions)
# recall <- recall_score(y_test, predictions)
# precision <- precision_score(y_test, predictions)
# f1 <- f1_score(y_test, predictions)
# balanced_accuracy <- balanced_accuracy_score(y_test, predictions)

# Display evaluation metrics
# print(roc_auc)
# print(avg_precision)
# print(accuracy)
# print(recall)
# print(precision)
# print(f1)
# print(balanced_accuracy)
```

```{r}
# Import XGBoost, as an exemplar, if this is the model you go for

# Try to load the xgboost package, install if necessary
# if (!requireNamespace("xgboost", quietly = TRUE)) {
#   install.packages("xgboost")
# }

# Load the xgboost package
# library(xgboost)

# Example of using XGBoost model
# Convert data to DMatrix, which is a recommended data structure for xgboost
# dtrain <- xgb.DMatrix(data = as.matrix(X_train), label = y_train)
# dtest <- xgb.DMatrix(data = as.matrix(X_test), label = y_test)

# Set parameters for XGBoost
# params <- list(
#   objective = "binary:logistic",
#   eval_metric = "auc"
# )

# Train the model
# model <- xgb.train(params, dtrain, nrounds = 100)

# Make predictions
# predictions <- predict(model, dtest)

# If you need class predictions, you can threshold the probabilities
# class_predictions <- ifelse(predictions > 0.5, 1, 0)

# Display predictions
# print(predictions)
# print(class_predictions)

```

Define your working directory and load the data.
```{r}
setwd('/Users/fredrikwhaug/code/bst-209-notebooks')
config <- config::get()
data_train<- read.csv(file.path(config$data_split, 'wids_train.csv'))
data_test <- read.csv(file.path(config$data_split, 'wids_test.csv'))
```

Let's look at the dimensions of our traning and test data.
```{r}
dim(data_train)
dim(data_test)
```


Define the outcome column
```{r}
outcome <- data$hospital_death
y_train <- data_train$outcome
y_test <- data_test$outcome
```

Select which features to use

Please comment the cells that do not apply to your track, and run the one that 
does apply. 

To make this model's results comparable with the one from the previous notebook, 
please make sure to keep the same features, except for the SpO2 and/or Lactate.

If you are Track 1 (modified SpO2 only).
We are selecting the minimum SpO2 because it is most representative of a 
patient's illness severity.
```{r}
features <- data[, c(
    'd1_spo2_min_new',
    'd1_lactate_max', #original
    'bmi',
    'age',
    'ethnicity',
    'gender'
    # same features as before
)]
```

If you are Track 2 (modified lactate only).
We are selecting the maximum lactate because it is most representative of a 
patient's illness severity.
```{r}
features <- data[, c(
    'd1_spo2_min', # original
    'd1_lactate_max_new',
    'bmi',
    'age',
    'ethnicity',
    'gender'
    # same features as before
)]

```

If you are Track 3 (both modified SpO2 and modified lactate)
```{r}
features <- data[, c(
    'd1_spo2_min_new',
    'd1_lactate_max_new',
    'bmi',
    'age',
    'ethnicity',
    'gender',
    # same features as before
)]
```

```{r}
X_train = data_train$features
X_test = data_test$features
```

Train a model
```{r}
# model <- ... # decide as a team which model you want to go with!
# model <- fit(model, X_train, y_train)
# predictions <- predict(model, X_test)
# prob_predictions <- predict(model, X_test, type = "prob")
```

Evaluate your model across race and ethnicity
```{r}
# first split the y_test per racial and ethnic group
# then for each group compute the evaluation metrics
# and you can compute overall metrics as well!
```